A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium} symbioses

### 2016-09-26 JTB_ms/JTB_ms.Rmd
more text

bibliography: library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes
---
```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load, include=FALSE}
# Source default parameters and functions
sapply(list.files(path="../../R", pattern="*.R", full.names=T), source, .GlobalEnv)
```

@@

| Symbol  | Description                            | Value  | Units              |
| ------- | -------------------------------------- | ------ | ------------------ |
| $n_{NH}$ | N:C molar ratio in host biomass | 0.19 | -- |
| $n_{NS}$ | N:C molar ratio in symbiont biomass | 0.2 | -- |
| $n_{NX}$ | N:C molar ratio in prey biomass | 0.13 | -- |
| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | $d^{-1}$ |
| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | $d^{-1}$ |
| $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | 0.9 | -- |
| $\sigma_{CH}$ | Proportion host carbon turnover recycled | 0.9 | -- |
| $\sigma_{NS}$ | Proportion symbiont nitrogen turnover recycled | 0.9 | -- |
| $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | 0.9 | -- |
| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | $molX \cdot CmolH^{-1} \cdot d^{-1}$ |
| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | $molX \cdot L^{-1}$ | 
| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | $molN \cdot CmolH^{-1} \cdot d^{-1}$ |
| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | $molN \cdot L^{-1}$ | 
| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | $molC \cdot CmolH^{-1} \cdot d^{-1}$ |
| $j_{HGm}$ | Maximum specific growth rate of host | 1 | $d^{-1}$ |
| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | $molC \cdot mol ph^{-1}$ |
| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | $m^2 \cdot CmolS^{-1}$ |
| $j_{NPQ}$ | Non-photochemical quenching capacity of symbiont | 40 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
| $k_{ROS}$ | Excess photon energy that doubles ROS prod. | 40-80 | $mol ph \cdot CmolS^{-1} \cdot d^{-1}$|
| $k$ | Exponent on ROS production rate | 1 | -- |
| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | $molC \cdot CmolS^{-1} \cdot d^{-1}$ |
| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | $d^{-1}$ |

@@

@@

To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and DIN, and plotted the specific growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B). Results are consistent with observations in corals: specific growth rates

![Steady state values of (A) specific growth ($Cmol \cdot Cmol^{-1} \cdot d^{-1}$) and (B) the symbiont to host biomass ratio ($CmolS \cdot CmolH^{-1}$) across gradients of external irradiance and dissolved inorganic nitrogen. Simulations for each combination of light and nutrients (21 points along each axis) were run for 100 days with a time step of 1 day. Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)

# Discussion  -  Potential applications/utility

>>>>

bibliography: library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes
---
```{r setup, include=FALSE}
library(reshape2); library(dplyr)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load, include=FALSE}
# Source default parameters and functions
sapply(list.files(path="../../R", pattern="^[^F].*.R", full.names=T), source, .GlobalEnv)
```

@@

| Symbol  | Description                            | Value  | Units              |
| ------- | -------------------------------------- | ------ | ------------------ |
| $n_{NH}$ | N:C molar ratio in host biomass | 0.19 | -- |
| $n_{NS}$ | N:C molar ratio in symbiont biomass | 0.2 | -- |
| $n_{NX}$ | N:C molar ratio in prey biomass | 0.13 | -- |
| $j_{HT}^0$ | Specific turnover rate of host biomass | 0.03 | d^-1^ |
| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | 0.03 | d^-1^ |
| $\sigma_{NH}$ | Proportion host nitrogen turnover recycled | 0.9 | -- |
| $\sigma_{CH}$ | Proportion host carbon turnover recycled | 0.9 | -- |
| $\sigma_{NS}$ | Proportion symbiont nitrogen turnover recycled | 0.9 | -- |
| $\sigma_{CS}$ | Proportion symbiont carbon turnover recycled | 0.9 | -- |
| $j_{Xm}$ | Maximum specific feeding rate of host | 0.1292 | molX CmolH^-1^ d^-1^ |
| $K_X$ | Half-saturation constant for prey uptake by host | 20e-6 | molX L^-1^ | 
| $j_{Nm}$ | Maximum specific DIN uptake rate by host | 0.048 | molN CmolH^-1^ d^-1^ |
| $K_N$ | Half-saturation constant for DIN uptake by host | 0.46e-6 | molN L^-1^ | 
| $j_{CO_2}^p$ | Passive CO~2~ delivery to symbiont | 4.04e-3 | molC CmolH^-1^ d^-1^ |
| $j_{CO_2}^a$ | Active CO~2~ delivery to symbiont | 0.32--18.0 | molC CmolH^-1^ d^-1^ |
| $j_{HGm}$ | Maximum specific growth rate of host | 1 | d^-1^ |
| $n_{LC}$ | Quantum yield of photosynthesis | 0.1 | molC mol ph^-1^ |
| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | 1.34 | m^2^ CmolS^-1^ |
| $j_{NPQ}$ | Non-photochemical quenching capacity of symbiont | 40 | mol ph CmolS^-1^ d^-1^ |
| $k_{ROS}$ | Excess photon energy that doubles ROS prod. | 40-80 | mol ph CmolS^-1^ d^-1^ |
| $k$ | Exponent on ROS production rate | 1 | -- |
| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | 2.8 | molC CmolS^-1^ d^-1^ |
| $j_{SGm}$ | Maximum specific growth rate of symbiont | 0.25 | d^-1^ |

@@

To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates

\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}

where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, all with units of Cmol Cmol^-1^ d^-1^.

@@

To analyze qualitative behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig. 2), which revealed patterns that are consistent with observed phenomena in corals.

_Specific growth rates_ are low at low light and DIN (~0.01-0.03 d^-1^), and begin increasing as both of these factors increase (Fig. 2A). Higher light alleviates light-limitation of photosynthesis (Fig. 2D), and consequently carbon-limitation of host growth (Fig. 2B). Similarly, elevated DIN alleviates nitrogen-limitation of host growth (Fig. 2B), a trend that was also predicted by the DEB model of @Muller:2009io, and has been observed experimentally [@MullerParker:1994uo; @Tanaka:2013dj]. Very high nutrient levels may have negative direct and indirect impacts on growth [@Shantz:2015dz], though these are not likely to occur until beyond the range of concentrations considered here [@FerrierPages:2000p1950]. The maximum specific growth rates of ~0.09 d^-1^, occurring between ~10-15 mol photons m^-2^ s^-1^ light and >1 µM DIN, are an order of magnitude higher than those typically observed for corals [@Osinga:2011kq]. The primary reasons for this overestimation of host growth are 1) the model does not account for calcification, and 2) it does not account for size-dependent reductions in growth that likely occur as individual polyps and/or colonies grow (Sebens1980?, probably better refs)... (i.e., the system continues growing exponentially in this model). However, the estimated specific growth rates are consistent with rates of biomass increase in *Aiptasia*, measured at 0.04 d^-1^ [@ArmozaZvuloni:2014ju], suggesting they are biologically realistic. They also match the values predicted by the full DEB model of @Muller2009:io. Furthermore, the relative differences across environmental gradients are likely to be the same.

As irradiance continues to increase above ~15-20 mol photons m^-2^ d^-1^, growth rates decline until the symbiosis is no longer viable above ~30 µmol photons m^-2^ d^-1^. The mechanism underlying this breakdown is the production of ROS caused by excitation energy in excess of what can be quenched by symbiont photosynthesis and non-photochemical quenching. ROS production has the consequence of increasing symbiont biomass loss (photodamage) and decreasing photosynthesis rates (photoinhibition), which have the net effect of reducing fixed carbon input to the system below levels necessary to sustain the symbiosis.

_Symbiont to host biomass ratios_ at low DIN (e.g. <1 µM) and high light (e.g., ~20 mol ph m^-2^ d^-2^) typical of coral reef environments are around ~0.1-0.3, which are consistent with values reported in the literature (@Muscatine:1981jy; @Edmunds:2011bv). S:H ratios increase under higher nutrients () and lower light ().

_

![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen. Simulations for each combination of light and nutrients (41 points along each axis) were run for 120 days with a time step of 1 day. Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig2.png)

# More things to show:
## varying light: consequences of excess light and light-limitation
## varying DIN: role of N-limitation and consequences of elevated N
## elevated nutrients cause system to become C-limited (photo: CO2 limited; host: fixed C limited)
## C-limitation increases susceptibility to bleaching
## consequences of a less efficient but more tolerant symbiont
## Rate of symbiont repopulation (Berner et al. 1993)
## Bicarbonate addition promotes coral growth (Marubini and Thake 1999)

### 2016-09-26 JTB_ms/JTB_ms.tex
more text

\(d^{-1}\)

@@

\(molX \cdot CmolH^{-1} \cdot d^{-1}\)

@@

\(molX \cdot L^{-1}\)

@@

\(molN \cdot CmolH^{-1} \cdot d^{-1}\)

@@

\(molN \cdot L^{-1}\)

@@

\(molC \cdot CmolH^{-1} \cdot d^{-1}\)

@@

\(molC \cdot mol ph^{-1}\)

@@

\(m^2 \cdot CmolS^{-1}\)

@@

\(mol ph \cdot CmolS^{-1} \cdot d^{-1}\)

@@

\(molC \cdot CmolS^{-1} \cdot d^{-1}\)

@@

@@

To analyze qualitative behavior, we ran the model to steady state across
gradients of external irradiance and DIN, and plotted the specific
growth rate (Fig. 2A) and symbiont to host biomass ratio (Fig. 2B).
Results are consistent with observations in corals: specific growth
rates

\begin{figure}[htbp]
\centering
\includegraphics{../img/Fig2.png}
\caption{Steady state values of (A) specific growth
(\(Cmol \cdot Cmol^{-1} \cdot d^{-1}\)) and (B) the symbiont to host
biomass ratio (\(CmolS \cdot CmolH^{-1}\)) across gradients of external
irradiance and dissolved inorganic nitrogen. Simulations for each
combination of light and nutrients (21 points along each axis) were run
for 100 days with a time step of 1 day. Negative steady state growth
rates, and corresponding S:H ratios, were set to zero.}
\end{figure}

\section{Discussion - Potential
applications/utility}\label{discussion---potential-applicationsutility}

@@

@@

@@

@@

>>>>

d\textsuperscript{-1}

@@

molX CmolH\textsuperscript{-1} d\textsuperscript{-1}

@@

molX L\textsuperscript{-1}

@@

molN CmolH\textsuperscript{-1} d\textsuperscript{-1}

@@

molN L\textsuperscript{-1}

@@

molC CmolH\textsuperscript{-1} d\textsuperscript{-1}

@@

molC mol ph\textsuperscript{-1}

@@

m\textsuperscript{2} CmolS\textsuperscript{-1}

@@

mol ph CmolS\textsuperscript{-1} d\textsuperscript{-1}

@@

molC CmolS\textsuperscript{-1} d\textsuperscript{-1}

@@

To aid in visualizing model results, we calculated values to indicate
the degree to which product formation at an SU was limited by
availability of either of its two substrates

\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}

where \(j_{S1}\) and \(j_{S2}\) are the specific input fluxes of the two
substrates and \(j_{Pm}\) is the maximum specific product formation
rate, all with units of Cmol Cmol\textsuperscript{-1}
d\textsuperscript{-1}.

@@

To analyze qualitative behavior, we ran the model to steady state across
gradients of external irradiance and nutrients (Fig. 2), which revealed
patterns that are consistent with observed phenomena in corals.

\emph{Specific growth rates} are low at low light and DIN
(\textasciitilde{}0.01-0.03 d\textsuperscript{-1}), and begin increasing
as both of these factors increase (Fig. 2A). Higher light alleviates
light-limitation of photosynthesis (Fig. 2D), and consequently
carbon-limitation of host growth (Fig. 2B). Similarly, elevated DIN
alleviates nitrogen-limitation of host growth (Fig. 2B), a trend that
was also predicted by the DEB model of Muller et al. (2009), and has
been observed experimentally (Muller-Parker et al. 1994; Tanaka et al.
2013). Very high nutrient levels may have negative direct and indirect
impacts on growth (\textbf{???}), though these are not likely to occur
until beyond the range of concentrations considered here (Ferrier-Pagès
et al. 2000). The maximum specific growth rates of \textasciitilde{}0.09
d\textsuperscript{-1}, occurring between \textasciitilde{}10-15 mol
photons m\textsuperscript{-2} s\textsuperscript{-1} light and
\textgreater{}1 µM DIN, are an order of magnitude higher than those
typically observed for corals (\textbf{???}). The primary reasons for
this overestimation of host growth are 1) the model does not account for
calcification, and 2) it does not account for size-dependent reductions
in growth that likely occur as individual polyps and/or colonies grow
(Sebens1980?, probably better refs)\ldots{} (i.e., the system continues
growing exponentially in this model). However, the estimated specific
growth rates are consistent with rates of biomass increase in
\emph{Aiptasia}, measured at 0.04 d\textsuperscript{-1} (\textbf{???}),
suggesting they are biologically realistic. They also match the values
predicted by the full DEB model of (\textbf{???}). Furthermore, the
relative differences across environmental gradients are likely to be the
same.

As irradiance continues to increase above \textasciitilde{}15-20 mol
photons m\textsuperscript{-2} d\textsuperscript{-1}, growth rates
decline until the symbiosis is no longer viable above
\textasciitilde{}30 µmol photons m\textsuperscript{-2}
d\textsuperscript{-1}. The mechanism underlying this breakdown is the
production of ROS caused by excitation energy in excess of what can be
quenched by symbiont photosynthesis and non-photochemical quenching. ROS
production has the consequence of increasing symbiont biomass loss
(photodamage) and decreasing photosynthesis rates (photoinhibition),
which have the net effect of reducing fixed carbon input to the system
below levels necessary to sustain the symbiosis.

\emph{Symbiont to host biomass ratios} at low DIN (e.g. \textless{}1 µM)
and high light (e.g., \textasciitilde{}20 mol ph m\textsuperscript{-2}
d\textsuperscript{-2}) typical of coral reef environments are around
\textasciitilde{}0.1-0.3, which are consistent with values reported in
the literature (Muscatine, R McCloskey, and E Marian (1981); Edmunds et
al. (2011)). S:H ratios increase under higher nutrients () and lower
light ().

\_

\begin{figure}[htbp]
\centering
\includegraphics{../img/Fig2.png}
\caption{Steady state values of \textbf{A)} specific growth (Cmol
Cmol\textsuperscript{-1} d\textsuperscript{-1}), \textbf{B)} the
symbiont to host biomass ratio (CmolS CmolH\textsuperscript{-1}),
\textbf{C)} relative C- or N-limitation of host biomass formation, and
\textbf{D)} relative CO\textsubscript{2}- or light-limitation of
symbiont photosynthesis, across gradients of external irradiance and
dissolved inorganic nitrogen. Simulations for each combination of light
and nutrients (41 points along each axis) were run for 120 days with a
time step of 1 day. Negative steady state growth rates, and
corresponding S:H ratios, were set to zero.}
\end{figure}

\section{More things to show:}\label{more-things-to-show}

\subsection{varying light: consequences of excess light and
light-limitation}\label{varying-light-consequences-of-excess-light-and-light-limitation}

\subsection{varying DIN: role of N-limitation and consequences of
elevated
N}\label{varying-din-role-of-n-limitation-and-consequences-of-elevated-n}

\subsection{elevated nutrients cause system to become C-limited (photo:
CO2 limited; host: fixed C
limited)}\label{elevated-nutrients-cause-system-to-become-c-limited-photo-co2-limited-host-fixed-c-limited}

\subsection{C-limitation increases susceptibility to
bleaching}\label{c-limitation-increases-susceptibility-to-bleaching}

\subsection{consequences of a less efficient but more tolerant
symbiont}\label{consequences-of-a-less-efficient-but-more-tolerant-symbiont}

\subsection{Rate of symbiont repopulation (Berner et al.
1993)}\label{rate-of-symbiont-repopulation-berner-et-al.-1993}

\subsection{Bicarbonate addition promotes coral growth (Marubini and
Thake
1999)}\label{bicarbonate-addition-promotes-coral-growth-marubini-and-thake-1999}

@@

\hypertarget{ref-Edmunds:2011bv}{}
Edmunds, Peter J, Hollie M Putnam, Roger M Nisbet, and Erik B Muller.
2011. ``Benchmarks in organism performance and their use in comparative
analyses.'' \emph{Oecologia} 167 (2): 379--90.
doi:\href{https://doi.org/10.1007/s00442-011-2004-2}{10.1007/s00442-011-2004-2}.

@@

\hypertarget{ref-FerrierPages:2000p1950}{}
Ferrier-Pagès, Christine, Jean-Pierre Gattuso, S Dallot, and Jean
Jaubert. 2000. ``Effect of nutrient enrichment on growth and
photosynthesis of the zooxanthellate coral Stylophora \ldots{}.''
\emph{Coral Reefs}.
\url{http://www.springerlink.com/index/3DRC3K1Q57TPC0UT.pdf}.

@@

\hypertarget{ref-MullerParker:1994uo}{}
Muller-Parker, Gisèle, Lawrence R McCloskey, Ove Hoegh-Guldberg, and PJ
McAuley. 1994. ``Effect of ammonium enrichment on animal and algal
biomass of the coral Pocillopora damicornis.'' \emph{Pacific Science} 48
(3). \url{http://scholarspace.manoa.hawaii.edu/handle/10125/2236}.

@@

\hypertarget{ref-Tanaka:2013dj}{}
Tanaka, Yasuaki, Akira Iguchi, Mayuri Inoue, Chiharu Mori, K Sakai,
Atsushi Suzuki, Hodaka Kawahata, and Takashi Nakamura. 2013. ``Marine
Pollution Bulletin.'' \emph{Marine Pollution Bulletin} 68 (1-2): 93--98.
doi:\href{https://doi.org/10.1016/j.marpolbul.2012.12.017}{10.1016/j.marpolbul.2012.12.017}.
